Foremost, I would like to express my sincere gratitude to my advisor Prof. Mausam for the continuous support of my Master's study and research, for his patience, motivation, enthusiasm, and immense knowledge. His guidance helped me in all the time of research and writing of this thesis. I could not have imagined having a better advisor and mentor for my Master's study. He consistently allowed me to frame my own work, but steered me in the right the direction whenever he thought I needed it.

I would also like to thank Dinesh Raghu for his immense support and guidance during the past two years. He brought structure to both my research and thinking process, allowing me to deep dive and grow into this field of study. I will cherish our late-night sprints during our paper submissions and our stimulating discussions regarding minute details of our algorithm or codebase. Dinesh consistantly brought out the best in me while giving me my own space to work and think, leading me to believe I have become a better researcher by the end of it.

I would also like to acknowledge my friend and labmate at IIT Delhi: Ayush Bhardwaj for being my right-hand man throughout all my course work and assignments, and I am gratefully indebted to his very valuable comments on this thesis. A heartful thanks to Aishwarya Sharma at IIT (BHU) Varanasi for lending her brilliant proofreading skills to my publications.

I thank Danish Contractor, Gaurav Pandey and Sachindra Joshi for their comments on my drafted publications. I extend gratitude to support my work received via IBM AI Horizons Network grant, an IBM SUR award, grants by Google, Bloomberg and 1MG, and a Visvesvaraya faculty award by Govt. of India. I thank Microsoft Azure sponsorships, and the IIT Delhi HPC facility for computational resources.s

Last but not the least, I would like to thank my family: my parentss for giving birth to me at the first place and supporting me spiritually throughout my life. They have taught me to always put my best effort in all my endevours and have consistanly set this example for me to follow. I am grateful to them for cultivating my curiousity and pushing me to be a good human being along my journey.